---
title: "Burn Index"
author: "cade"
date: "11/17/2018"
output: html_notebook
---
### Description 

This code notebook has processes Landsat 8 files that are downloaded from Earth Explorer in the following order:
1. Untars files
2. Layer Stacks Multi-spectral bands
3. Calculates NBR
4. Calculate dNBR
5. Applies a severity index

Future:
6. Crops file to the burn perimeter 
7. Totals the acreage for each burn type specified by the burn severity. 

Future updates:
All loops will be switched to functions to improve efficiency. 

** USER REQUIREMENTS **
Downloaded Landsat 5 or 8 files. 
A CSV with the following information

### Code

```{r}
#packages
require(raster)
require(rgdal)
require(stringr)
require(RStoolbox)
require(tictoc)
require(readxl)
require(tidyverse)
require(lubridate)
require(tools)
require(magrittr)

```

**User Defined Variables**
*Note: each user must edit this portion*
Any line code with the symbol #~! will be updated in the future. The type of update will be listed in the code block.

```{r}
# userDefined variables
# 1) burn Info sheet 
# must have columns 
# Name,	County,	Acres,	Start date,	Containment date,	lat,	long,	PreDate,	PostDate,	Sensor,	path,	row	notes,	cadeComment,	cloudMask,	MultipleScenes
burnInfoDir <- "Z:\\Cade\\CalFire\\Data\\sampleHistorical\\sampleFires_nov23_cade.xlsx"
# 2) Image Directory - tar.gz files for all of the acquired sceness
imgDir <- "./Data/Sample_fires"
# 3) Output Directory of images that are layer stacked
outSR <- "./Output/StackedSR_historical"
# 4) Output Directory of BurnIndex
outBurn <- "./Output/NBR_historical"
# 5) Output Directory of dNBR
outDNBR <- "./Output/dNBR_historical"
# 6) Output Directory of burnSeverity dNBR
# not ued yet 
# outSev <- "Z:\\Cade\\CalFire\\Output\\Severity"
# 7) input  of fire perimeter 
inShp <- "./Data/shapefiles/perimeter"
# 8) Output of table for acres
outArea <- "./Output/AreaCalcs"


```

```{r}
# edit burn info data to have the correct pre and post fire information
# Thi should have also a more elegant way to add 0s to path and row - if there are three numbers do nothing 
# if there are 2 numbers then add 0s
burnInfo <- read_excel(burnInfoDir) %>%
  #filter(cadeComment == "good") %>%
  mutate(preImg = paste0(outBurn,"/",Sensor,paste0(0,path),paste0(0,row),year(PreDate),
                         substr(PreDate,6,7),substr(PreDate,9,10),"01T1","_nbr.tif"),
         postImg = paste0(outBurn,"/",Sensor,paste0(0,path),paste0(0,row),year(PostDate),
                          substr(PostDate,6,7),substr(PostDate,9,10),"01T1","_nbr.tif"),
         dNBRImg = paste0(outDNBR,"/",Name,"_",`Start date`, "_", Sensor, paste0(0,path),
                          paste0(0,row),"_dNBR.tif"))
csvOutName = paste0(file_path_sans_ext(burnInfoDir),"_updated.csv")
write_csv(burnInfo,csvOutName)
```


### Step 1: Selection of Images 
All Images were selected using the USGS Earth Explorer webpage

### Step 2: Download Data
Surface reflectance products that have been ordered using Earth Explorer can be downloaded using espa.tools. The username, password and order ID number must be known
```{r}
# example
#require(espa.tools)
# earthexplorer_download(usgs_eros_username = "cade", usgs_eros_password = "1234", 
#                        output_folder = "Z:\\Data\\SurfaceReflctance",
#                        ordernum = "espa-C.ade92@gmail.com-0101809057908")


```

### Step 3: Unzip files
This will result in a number of  tar.gz files that have a single GEOtiff for each band and some ancillary information. The following step untars the file and creates a new out folder based on the file name (see comments in ```landUntar.R``` for further details). 
```{r}
# function for untarring and organizing landsat level 2 data
source("./Functions/landUntar.R")
#list original tared data
orig.data = list.files(imgDir,pattern=".tar$|.tar.gz$",full.names = T)
tic("unzip")
lapply(orig.data,landUntar)
toc()
## 2255.02 
## Ran Wednesday 11/13 cade computer 

```


#### Step 4: Layerstack bands 
Stack the bands in each folder.
You want the sr_band1.tif - sr_band7.tif

To update in the future, will include a quick looks with the fire point and zoom in of the area. 

```{r}
# list all your directories
folders <-list.dirs(path = imgDir, full.names = TRUE, recursive = FALSE)[6]

source("./Functions/stackLandsat.R")
tic("stack")
lapply(folders,stackLandsat, outSF = outSR)
toc()

# re-ran Friday 11/16 cade computer
```

#### Step 5: Calculate NBR

```{r}
# list image stacks 
imgSF <- list.files(outSR,  pattern = "tif$", full.names = T)
tic("calcualte NBR")
source("./Functions/calcNBR.R")
lapply(imgSF,calcNBR, outDir = outBurn) 
toc()

# 599.58 

```

### Step 6: Difference burn images

```{r}

beginCluster(n=5)
tic("calcualte dNBR")
for (i in 2:nrow(burnInfo)){
  # subset data frame by row
  fireEvent <- burnInfo %>%
    slice(i) #i
  # read in preFire ndr
  preFireNBR <- raster(fireEvent$preImg)
  # read in post fire ndr
  postFireNBR <- raster(fireEvent$postImg)
  # calculate nbr
  dNBR <- preFireNBR - postFireNBR
  writeRaster(dNBR, fireEvent$dNBRImg, format ="GTiff", overwrite = T)
  rm(preFireNBR)
  rm(postFireNBR)
  rm(dNBR)
  removeTmpFiles(h=0)
}
endCluster()
toc

```



### Step 7 : Calculate burn severity
This part creates rasters and computes area

** Severity index 1**
Christiana modified severity index from 
https://www.earthdatascience.org/courses/earth-analytics/multispectral-remote-sensing-modis/calculate-dNBR-R-Landsat/ 
```{r}
# Create a reclasss file 
# we can change this at any time 
sev1 <- c(-Inf,-0.1, 1, # enhanced regrowth
                -0.1,0.1,2, # unburned
                0.1,0.27,3, # low severity
                0.27,.66,4, # moderate severity
                .66,1.3,5, # moderate to high severity
                1.3, Inf,6) # high severity
# Create classs matrix 
sev1_m <- matrix(sev1, 
                    ncol = 3,
                    byrow = T)
# key
sev1_Key <- tibble( value = c(1:6),
                    severity = c("enhanced regrowth","unburned","low severity","moderate severity","moderate to high","high severity"))

```

** Severity index 2**
http://gsp.humboldt.edu/olm_2015/Courses/GSP_216_Online/lesson5-1/NBR.html
```{r}
# Create a reclasss file 
# ~! can create this into a fucntion?
# we can change this at any time 
sev2 <- c(-Inf,-0.1, 1, # enhanced regrowth
                -0.1,0.1,2, # unburned
                0.1,0.27,3, # low severity
                0.27,.66,4, # moderate severity
                .66,Inf,5) # # high severity
# Create classs matrix 
sev2_m <- matrix(sev2, 
                    ncol = 3,
                    byrow = T)
# create key
sev2_Key <- tibble( value = c(1:5),
                    severity = c("enhanced regrowth","unburned","low severity","moderate severity","high severity"))
```

** Severity index 2**
http://gsp.humboldt.edu/olm_2015/Courses/GSP_216_Online/lesson5-1/NBR.html
```{r}
# Create a reclasss file 
# ~! can create this into a fucntion?
# we can change this at any time 
sev2 <- c(-Inf,-0.1, 1, # enhanced regrowth
                -0.1,0.1,2, # unburned
                0.1,0.27,3, # low severity
                0.27,.66,4, # moderate severity
                .66,Inf,5) # # high severity
# Create classs matrix 
sev2_m <- matrix(sev2, 
                    ncol = 3,
                    byrow = T)
# create key
sev2_Key <- tibble( value = c(1:5),
                    severity = c("enhanced regrowth","unburned","low severity","moderate severity","high severity"))
```


```{r}
# read in function
source("./Functions/severityIndex.R")
# list dNbr files
list.dNbr <- list.files(outDNBR, pattern = "_dNBR.*tif$",full.names = T)

## apply first severity index
lapply(list.dNbr, severityIndex, reclass_m = sev1_m, appendName = "sev1")

## apply sectiond index
lapply(list.dNbr, severityIndex, reclass_m = sev2_m, appendName = "sev2")


```


```{r}
# ~! should make more systematic
# probably just take from the table?
# this will need to be updated post crop
list.sev <- list.files(outDNBR, pattern = "sev", full.names = T)
# list shapefiles (need to fix)
listShp <- list.files(inShp,pattern = "shp$", full.names = T, recursive =  T)

for(i in 1:length(list.sev)){
  rasterFileName <- list.sev[i] #i
  # fire event
  fireEvent <- tolower(str_extract(basename(rasterFileName),"(Gap|Paradise|Soberanes)"))
  for (j in 1:length(listShp)){
    shpName <- listShp[j] # j
  # ~! bad string matching
    shpFire <- str_extract(basename(shpName),"(gap|paradise|soberanes)")
    if(fireEvent == shpFire){
             # stack raster
    myRaster <- stack(rasterFileName)
    # read in fire perimeter
    roi <- readOGR(shpName) 
    # reproject perimeter to raster
    roi_trans <- spTransform(roi, CRS = crs(myRaster))
    # crop raster
    rast.crop <- mask(crop(myRaster, roi_trans),roi_trans) 
    # write Crop raster for mapping 
    writeRaster(rast.crop, paste0(file_path_sans_ext(rasterFileName),"_crop"),format = "GTiff",overwrite = T)
    # satellite res
    resSat<- res(rast.crop)
    # calcualte area by raster pixel value
    areaCalc <- as.data.frame(freq(rast.crop, useNA="no")) %>%
      setNames(c("value","count")) %>% 
      mutate(area_km2 = count * prod(resSat) *  1e-06,
             area_acre = area_km2 * 247.105,
             productName = basename(rasterFileName),
             sevIndex = str_extract(productName, "(sev1|sev2)"),
             key = ifelse(sevIndex == "sev2",sev2_Key$severity,sev1_Key$severity))
    write_csv(areaCalc, paste0(outArea,"/",basename(rasterFileName),"_area.csv"))


    } else {
      print("no match")
    }
  }
}



################# Turn the above into a function somehow? ##################
# cropAreaCount <- function(rasterFileName, roiName ){
#   
#   rasterFileName <- list.sev[1]
#   # stack raster
#   myRaster <- stack(rasterFileName)
#   # read in fire perimeter
#   roi <- readOGR(listShp[2])
#   # reproject perimeter to raster
#   roi_trans <- spTransform(roi, CRS = crs(myRaster))
#   # crop raster
#   rast.crop <- mask(crop(myRaster, roi_trans),roi_trans) 
#   #
#   # satellite res
#   resSat<- res(rast.crop)[[1]]
#   areaCalc <- as.data.frame(freq(rast.crop, useNA="no")) %>%
#     setNames(c("value","count")) %>% 
#     mutate(area_km2 = count * prod(resSat) *  1e-06,
#            area_acre = area_km2 * 247.105)
#  }

```


